\textbf{Species and genus plant classification with NEON hyperspectral remote sensing data}

### 2020-09-10 analysis/paper/paper.Rmd
adding Max's gist code for extracting the spectral data
added outline for paper

```{r, results='hide', message=FALSE, include=FALSE}
# Run the main analysis
```

@@

@@

# Methods {-}

@@

@@

@@

@@

>>>>

```{r, results='hide', message=FALSE, include=FALSE}
# Extract hyperspectral reflectance data for each individual plant
#R/spectral-lib.Rmd

# Run the main analysis

```

@@

1 - motivation for work, why bother?
  - unprecedented access to unique hyperspectral airborne imagery from NEON (cite)
  - acknowledge additional data to add info (lidar, rgb) (cite) 
  - how meaningful can the HS data alone be to predict species? accuracies? 
  - we expect there to be meaningful information in the HS data for species/genus
2 - knowledge gaps in lit/field?
  - how useful are HS data across many species, ecosystems, geographies? 
  - where is the most important information contained? which wavelengths most important for species discrimination? differ across sites? (cite) 
  - generally more site-specific studies? 
  - prediction of species vs genus? How much easier is genus? Or is this difficult too? (cite hierarchical taxonomic classification recent study)
3 - how we addressed the knowledge gaps
  - the need to generalize methods across locations? --> NEON
  - which wavelengths are most important
  - data harmonization across NEON sites (why are the wavelengths different? Ask NEON person?)

@@

# Methods {-}
NEON data collection protocols
AOP flights
Plant mapping 
Network of sites and observations (how many? ecoclimatic domains? ecosystem coverage?

Figure - NEON sites and domains, which sites do we have data from? # individuals from each site? 
Acknowledge imbalance spatially and temporally...
Any sites not represented? Like Hawaii

## Data extraction {-}

Extracted hyperspectral data for all mapped stems at NEON sites
Assumed plant locations don't change over time 
Filtered plants to keep only "live" ones
NEON data products (cite)
Spectrometer orthorectified surface directional reflectance mosaic
Woody plant vegetation structure field based measurements
Mask created for water absorption bands (cite)
Packages neonUtilities, geoNEON, neonhs (cite)

@@

? Filtered taxonomic levels, labels species & genus
Wavelength interpolation - filtered out the PUUM site in Hawaii because of wavelengths
Filtering based on # samples
Train/validate split 80/20

Figure - histogram of species counts, show which ones were used to train
Figure - cleaned data, make it legible

@@

Random forest classifier
Hyperparameter optimization (mtry, ntree)
Overall accuracy, confusion matrices, variable importance

@@

Classification accuracy was higher at the genus level (60%) compared to the species level (55%)

Figure - confusion matrices
Figure - variable importance vs wavelength

@@

Restate findings at high level

How do our findings compare to other similar studies. Consistent? Differences? (cite IDTReeS) Other ranges of accuracy using NEON HS data? How many sites too?

Potential limitations
  - class imbalance - we tried to filter based on # samples, but our model has then only seen common species and would be unable to identify rare ones
  - filtering shadows, non-vegetation pixels.... (cite a paper from last IDTReeS comp)

Future directions, next steps
  - targeted species/genus for increased sampling to improve representation in the data set?
   - more advanced models. ML DL
   - RF treats each feature as independent
   - spectra are sequences of wavelengths - explore DL methods for sequences (i.e. recurrent neural networks)
   - hierarchical taxonomic approach - borrow information from closely related species (i.e. common and rare species > use common species observations)
   - (cite our other paper) intelligent crown geometry processing 
   - (ben weinstein) tree crown detection - extract multiple pixels per individual using rectangular crown boundaries
   - herbarium / hyperspectral data measurements at other scales. handheld spectrometer. other spectral libraries. 
   - combine HS with RGB, lidar

We conclude that HS data allow you to get __ accuracy
Hopes for outcomes

# Appendix / Supplement ideas {-}

- Species histogram to show which ones were excluded for few # samples 
- Wavelength interpolation figures?
